# (PART) Module 03 {-}

```{r include = FALSE}
source("common.R")
# install.packages("devtools")
if (!require("tweetrmd"))  devtools::install_github("gadenbuie/tweetrmd")
library(tweetrmd) #... embedding tweets
```

```{r links, child="links.md"}
```

# Welcome to Ethics in Testing

This module is covers Ethics in Testing. Please watch the videos and work your way through the notes. **The videos start on the next page.**  You can find the module playlist [here](https://www.youtube.com/playlist?list=PLKrrdtYgOUYb-wsx3rHZI5e33cE1sLE9R). 

## Learning Goals

By the end of this module, you should be able to achieve all the learning goals outlined below.

- Discuss how the profession of psychology regulates itself internally and cite examples of documents that the APA has promulgated for the use of psychologists
- Explain the major reasons why the use of psychological tests needs to be restricted to those who are competent in their use
- Describe how you would gauge whether someone is a qualified psychologist
- List three examples of practices concerning the distribution and marketing of psychological tests that are deemed unacceptable and explain why they might be harmful
- Discuss how the purpose of which testing is conducted affects the issue of protecting the privacy of test takers
- Explain what is meant by "informed consent" as fully as possible
- Discuss the issues of access to test results from the standpoint of test-takers parents third persons and institutions
- Identify and discuss the major precautions that should apply to the communication of test results to the individual test taker
- Describe the major legal regulations in the area of employee selection procedures
- Cite an example from the field of employment testing of a well-intentioned effort to do away with barriers to equal opportunity that has proved to be controversial
- Discuss the issue of relevant vs. irrelevant cultural factors as it applies to the content of psychological tests
- Describe the rationale used by those who view standardized testing as having a positive function with regard to protecting the civil rights of minorities and women

## Module Materials

### Explore
* Readings
	* Located on Canvas

* Videos
  * Located in the subchapters of this module
* Slidedecks
	* Located on Canvas
	
### Engage
* Activities
	* Located on Canvas
	
### Evaluate
* Quiz
	* Located on Canvas



### Estimated Video Length

<!--https://www.youtube.com/playlist?list=PLKrrdtYgOUYao_7t5ycK4KDXNKaY-ECup & https://ytplaylist-len.herokuapp.com/ -->

No of videos : 7

Average length of video : 16 minutes, 59 seconds

Total length of playlist : 1 hour, 58 minutes, 55 seconds

- At 1.25x : 1 hour, 35 minutes, 8 seconds
- At 1.50x : 1 hour, 19 minutes, 16 seconds

# Lecture: Ethics in Testing

##  Ethics, Equity, and APA guidelines (03.01)


<!-- not yet updated -->

```{r, echo=FALSE}
embed_url("https://youtu.be/eFT2JgPKk7U") %>%
  use_align("center")
```


## Responsibilities of Publishers (03.02)


<!-- not yet updated -->

```{r, echo=FALSE}
embed_url("https://youtu.be/y8Eg6gS1WT8") %>%
  use_align("center")
```

## Privacy and Test User Responsibilities (03.03)

<!-- not yet updated -->

```{r, echo=FALSE}
embed_url("https://youtu.be/5xAvbRH_pWE") %>%
  use_align("center")
```

## Considerations for Individual Differences (03.04)


<!-- not yet updated -->

```{r, echo=FALSE}
"https://youtu.be/nZi2ykCATnE" %>%
embed_url() %>%
  use_align("center")
```


## Legal Considerations (03.05)

<!-- not yet updated -->

```{r, echo=FALSE}
"https://youtu.be/lzUU3CHXZEo" %>%
embed_url() %>%
  use_align("center")
```

## Ethics from the Test-Taker's Perspective (03.06)
<!-- not yet updated -->
```{r, echo=FALSE}
embed_url("https://youtu.be/HnZ_HsvDNxM") %>%
  use_align("center")
```

## Potential Problems from the Test-Taker's Perspective (03.07)

<!-- not yet updated -->
```{r, echo=FALSE}
embed_url("https://youtu.be/2IR-C_ZG_dA") %>%
  use_align("center")
```
