# (PART) Module 04 {-}

```{r include = FALSE}
source("common.R")
# install.packages("devtools")
if (!require("tweetrmd"))  devtools::install_github("gadenbuie/tweetrmd")
library(tweetrmd) #... embedding tweets
```

```{r links, child="links.md"}
```

# Welcome to Norms and Meanings of Test Scores

This module is covers Norms and Meanings of Test Scores. Please watch the videos and work your way through the notes. **The videos start on the next page.**  You can find the module playlist [here](https://www.youtube.com/playlist?list=PLKrrdtYgOUYbKLQBwbeSVN4mp0DPkNG6t). 

## Learning Goals

By the end of this module, you should be able to achieve all the learning goals outlined below.



## Module Materials

### Explore
* Readings
	* Located on Canvas

* Videos
  * Located in the subchapters of this module
* Slidedecks
	* Located on Canvas
	
### Engage
* Activities
	* Located on Canvas
	
### Evaluate
* Quiz
	* Located on Canvas



### Estimated Video Length

<!--https://www.youtube.com/playlist?list=PLKrrdtYgOUYao_7t5ycK4KDXNKaY-ECup & https://ytplaylist-len.herokuapp.com/ -->



# Lecture: Norms and Meaning of Test Scores

## Norms and Meaning of Test Scores (04.01)


<!-- not yet updated -->

```{r, echo=FALSE}
"https://www.youtube.com/watch?v=WcoBYrwKQJo" %>%
 embed_url() %>%
  use_align("center")
```


## Statistics as Summary (04.02)


<!-- not yet updated -->

```{r, echo=FALSE}
"https://www.youtube.com/watch?v=d4AIfgUdEGY" %>%
 embed_url() %>%
  use_align("center")
```

## Measures of Spread (04.03)

<!-- not yet updated -->

```{r, echo=FALSE}
"https://youtu.be/oilUDTYEBVA" %>%
 embed_url() %>%
  use_align("center")
```

## Transforming Raw Scores (04.04)


<!-- not yet updated -->

```{r, echo=FALSE}
"https://www.youtube.com/watch?v=v1diShA7Nug" %>%
 embed_url() %>%
  use_align("center")
```


## Developmental Norms (04.05)

<!-- not yet updated -->

```{r, echo=FALSE}
"https://www.youtube.com/watch?v=9q--3XpVwq0" %>%
 embed_url() %>%
  use_align("center")
```

## Within-Group Norms (04.06)
<!-- not yet updated -->
```{r, echo=FALSE}
"https://www.youtube.com/watch?v=QBBXrIztZ64" %>%
 embed_url() %>%
  use_align("center")
```

## Referenced Tests Types (04.07)

<!-- not yet updated -->
```{r, echo=FALSE}
"https://www.youtube.com/watch?v=Q4BjtRh9JDA" %>%
 embed_url() %>%
  use_align("center")
```

## Minimums, Cutoffs, and Expectancy Tables (04.08)

<!-- not yet updated -->
```{r, echo=FALSE}
"https://www.youtube.com/watch?v=zCAj7EaVBWM" %>%
 embed_url() %>%
  use_align("center")
```

## Interpreting Test Scores Realistically: Remastered (04.09)

I remastered this video on Interpreting Test Scores Realistically. It was originally produced in 1961. The original can be found [here](https://archive.org/details/interpretingtestscoresrealistically)

<!-- not yet updated -->
```{r, echo=FALSE}
"https://www.youtube.com/watch?v=jlk32-Tmt80" %>%
 embed_url() %>%
  use_align("center")
```
