# (PART) Other Coolness {-}

```{r include = FALSE}
source("common.R")

# install.packages("devtools")

if (!require("tweetrmd"))  devtools::install_github("gadenbuie/tweetrmd")
library(tweetrmd) #... embedding tweets
library(vembedr)  #... embedding youtube videos
library(knitr)
```

```{r links, child="links.md"}
```

# Good Resources

## Learning R
* https://psychnerdjae.github.io/into-the-tidyverse/
* [Automatic Grading with RMarkdown example](https://github.com/rundel/hw01)
* [Git/Github for virtual learning](https://afredston.github.io/learn-git/learn-git.html) (from [this tweet](https://twitter.com/AFredston/status/1354865931772682243))
* [Learn-Datascience-for-Free](https://github.com/therealsreehari/Learn-Datascience-for-Free)
* https://allisonhorst.shinyapps.io/dplyr-learnr/

## Psychology Scales

* [International Personality Item Pool](https://ipip.ori.org/)
* [Open-Source Psychometrics Project](https://openpsychometrics.org/)
* [APA PsycTests](https://www.apa.org/pubs/databases/psyctests)
* [Measurement Instrument Database for the Social Sciences](https://www.midss.org/)
* [Paul Spector’s assessment archive](https://paulspector.com/assessments/assessment-archive/)
* [GESIS Open Access Repository for Measurement Instruments](https://zis.gesis.org/en)
* [REDCap's library of scales](https://redcap.vanderbilt.edu/consortium/library/search.php)
* [PhenX Toolkit](https://www.phenxtoolkit.org/)
* [Registry of Scales and Measures: Psychological Tests, Scales, Questionnaires and Checklists](http://www.scalesandmeasures.net/)
